---
title: "Decision Tree"
output: html_document
---

### .We will be using the iris data set for this example.
```{r}
head(iris)
str(iris)
```

### We will build a decision tree for the iris data set. We will be using the “ctree” function. The first parameter will be the formula we want to use. Here we will choose “Species” as the dependent variable and “Sepal.Length”, “Sepal.Width”, “Petal.Length” and “Petal.Width” will be the independent variables.

```{r}
install.packages("party")
library(party)
library(plyr)
library(readr)


iris_ctree <- ctree(Species~ Sepal.Length + Sepal.Width +Petal.Length  + Petal.Width , data=iris)
print(iris_ctree)
```
### Now, let's plot the decision tree

```{r}
plot(iris_ctree , type="simple")

```

### Decision trees are also used as classifiers. To this aim, let's split the iris dataset into training and test set.

```{r}
train_index_iris <- sample(1:nrow(iris), 0.7 *nrow(iris))
train_set_iris <- iris[train_index_iris, ]
test_set_iris <- iris[-train_index_iris,]

```

### Let's run our model on the training set

```{r}
iris_ctree_model <- ctree(Species~ Sepal.Length + Sepal.Width +Petal.Length  + Petal.Width , data=train_set_iris )

iris_ctree_model

```
### Now, let's make our prediction on the test set

```{r}
iris_ctree_prediction <- predict(iris_ctree_model , test_set_iris)
head(iris_ctree_prediction)

```
### Let's see the confusion matrix.

```{r}
table(iris_ctree_prediction, test_set_iris$Species)

```










