---
title: 'CMTH 642 Data Analytics: Advanced Methods'
output:
  html_document: default
  pdf_document: default
  word_document: default
---
<center> <h1> Assignment 1 (10%) </h1> </center>
<center>  <h3>  </h2> </center>
<center> <h3> </h2> </center>
---

#### 1. Read the csv files in the folder. (4 points)
```{r}
# INSERT YOUR ANSWER HERE

Location <- getwd()

Macro_Nut <- read.csv(paste(Location, "USDA_Macronutrients.csv", sep = "/"))

Micro_Nut <- read.csv(paste(Location, "USDA_Micronutrients.csv", sep = "/"))


```

#### 2. Merge the data frames using the variable "ID". Name the Merged Data Frame "USDA". (4 points)
```{r}
# INSERT YOUR ANSWER HERE

Merged_USDA <- merge(Macro_Nut, Micro_Nut, by="ID")


```

#### 3. Check the datatypes of the attributes. Delete the commas in the Sodium and Potasium records. Assign Sodium and Potasium as numeric data types. (4 points)
```{r}
# INSERT YOUR ANSWER HERE
str(Merged_USDA)

Merged_USDA$Sodium = gsub(",","",Merged_USDA$Sodium, fixed = TRUE )
Merged_USDA$Sodium = as.numeric(Merged_USDA$Sodium , fixed = TRUE)

Merged_USDA$Potassium = gsub(",","",Merged_USDA$Potassium)
Merged_USDA$Potassium = as.numeric(Merged_USDA$Potassium)

#Applying sapply function to check datatypes
sapply(Merged_USDA, class)

```

```{r}
head(Merged_USDA)

```

```{r}
tail(Merged_USDA)
```




#### 4. Remove records (rows) with missing values in more than 4 attributes (columns). How many records remain in the data frame? (4 points)


```{r}

Original_Record<- cat("Original_Record: ", nrow(Merged_USDA))


Merged_USDA <- Merged_USDA[rowSums(is.na(Merged_USDA)) < 5,]


Remaning_Record <- cat("Remaning_Record: ", nrow(Merged_USDA))




```




#### 5. For records with missing values for Sugar, Vitamin E and Vitamin D, replace missing values with mean value for the respective variable. (4 points)
```{r}
# INSERT YOUR ANSWER HERE


Merged_USDA$Sugar[is.na(Merged_USDA$Sugar)] = mean(Merged_USDA$Sugar[!is.na(Merged_USDA$Sugar)])

Merged_USDA$VitaminE[is.na(Merged_USDA$VitaminE)] = mean(Merged_USDA$VitaminE[!is.na(Merged_USDA$VitaminE)])

Merged_USDA$VitaminD[is.na(Merged_USDA$VitaminD)] = mean(Merged_USDA$VitaminD[!is.na(Merged_USDA$VitaminD)])

#Validate our function
summary(is.na(Merged_USDA$Sugar))
summary(is.na(Merged_USDA$VitaminE))
summary(is.na(Merged_USDA$VitaminD))

```

#### 6. With a single line of code, remove all remaining records with missing values. Name the new Data Frame "USDAclean". How many records remain in the data frame? (5 points)
```{r}
# INSERT YOUR ANSWER HERE

USDAclean <- Merged_USDA[complete.cases(Merged_USDA),] 
cat("Number of remaining records:", nrow(USDAclean))




```

#### 7. Which food has the highest sodium level? (5 points)
```{r}
# INSERT YOUR ANSWER HERE


USDAclean$Description[USDAclean$Sodium == max(USDAclean$Sodium)]

```

#### 8. Create a histogram of Vitamin C distribution in foods. (5 points)
```{r}
# INSERT YOUR ANSWER HERE


hist(USDAclean$VitaminC,xlab = "Vitamin C" ,main="Histogram of Vitamin C",xlim=c(0,100), breaks=100 , col = c("red","green","blue", "yellow"))

```

#### 9. Create a boxplot to illustrate the distribution of values for TotalFat, Protein and Carbohydrate. (5 points)
```{r}
# INSERT YOUR ANSWER HERE

boxplot(USDAclean$TotalFat, USDAclean$Protein, USDAclean$Carbohydrate ,main = "Boxplot of TotalFat, Protein and Carbohydrate", xlab="TotalFat, Protein and Carbohydrate",col= c("pink","green", "yellow"))

```

#### 10. Create a scatterplot to illustrate the relationship between a food's TotalFat content and its Calorie content. (5 points)
```{r}
# INSERT YOUR ANSWER HERE

plot(USDAclean$TotalFat, USDAclean$Calories, col=c("purple","orange"), xlab = "TotalFat", ylab = "Calories", main = "Totalfat Vs Calories")


```

#### 11. Add a variable to the data frame that takes value 1 if the food has higher sodium than average, 0 otherwise. Call this variable HighSodium. Do the same for High Calories, High Protein, High Sugar, and High Fat. How many foods have both high sodium and high fat? (5 points)
```{r}
# INSERT YOUR ANSWER HERE
USDAclean$HighSodium <- 0
USDAclean$HighSodium[USDAclean$Sodium > mean(USDAclean$Sodium)]=1 

USDAclean$HighCalories <- 0
USDAclean$HighCalories[USDAclean$Calories > mean(USDAclean$Calories)]=1

USDAclean$HighProtein <- 0
USDAclean$HighProtein[USDAclean$Protein > mean(USDAclean$Protein)]=1

USDAclean$HighSugar <- 0
USDAclean$HighSugar[USDAclean$Sugar > mean(USDAclean$Sugar)]=1

USDAclean$HighFat <- 0
USDAclean$HighFat[USDAclean$TotalFat > mean(USDAclean$TotalFat)]=1


cat(sum(apply(USDAclean[c("HighSodium", "HighFat")], 1, function(x) sum(x) == 2)),
"foods have both high sodium and high fat.")

```

#### 12. Calculate the average amount of iron, for high and low protein foods. (5 points)
```{r}
# INSERT YOUR ANSWER HERE


tapply(USDAclean$Iron, USDAclean$HighProtein, mean)

```

#### 13. Create a function for a "HealthCheck" program to detect unhealthy foods. Use the algorithm flowchart below as a basis. (5 points)
```{r}
require(jpeg)
img<-readJPEG("HealthCheck.jpg")
plot(1:4, ty = 'n', ann = F, xaxt = 'n', yaxt = 'n')
rasterImage(img,1,1,4,4)
```
```{r}
# INSERT YOUR ANSWER HERE

HealthCheck = function(x){
if(x$HighSodium == 0) return("Pass")
else if(x$HighSugar == 0)return("Pass")
else if(x$HighFat == 0 )return("Pass")
else return("Fail")
}


```

#### 14. Add a new variable called HealthCheck to the data frame using the output of the function. (5 points)
```{r}
# INSERT YOUR ANSWER HERE

for (i  in 1:nrow(USDAclean)) {
  USDAclean$HealthCheck[i]=HealthCheck(USDAclean[i,])
  
}
str(USDAclean)

```

#### 15. How many foods in the USDAclean data frame fail the HealthCheck? (5 points)
```{r}
# INSERT YOUR ANSWER HERE
cat("The number is: ",sum(USDAclean$HealthCheck == "Fail"))

```

#### 16. Visualize the correlation among Calories, Protein, Total Fat, Carbohydrate, Sodium and Cholesterol. (5 points)
```{r}
# INSERT YOUR CODE HERE.
# First for more investigation,  boxplots are drawn for each variable.
boxplot(USDAclean$Calories, ylab="Amount",xlab="Calories" ,col="yellow")

```

```{r}
boxplot(USDAclean$Protein, ylab="Amount", xlab="Protein" , col="Beige")
```

```{r}
boxplot(USDAclean$TotalFat, ylab="Amount", xlab="TotalFat",col="light Blue")
```

```{r}
boxplot(USDAclean$Carbohydrate, ylab="Amount",xlab="Carbohydrate", col="light Pink")
```

```{r}
boxplot(USDAclean$Sodium, ylab="Amount", col="Purple" , xlab="Sodium")
```

```{r}
boxplot(USDAclean$Cholesterol, ylab="Amount", col="Orange", xlab="Cholesterol")


```


```{r}
# Then I Use the Pearson Correlation to compute the correlation.
Selected_Features <- USDAclean[,c('Calories', 'Carbohydrate',
'Protein', 'TotalFat', 'Sodium')]

#install.packages("PerformanceAnalytics")
library(PerformanceAnalytics)

chart.Correlation(Selected_Features, histogram = TRUE)

```



#### 17. Is the correlation between Calories and Total Fat statistically significant? Why? (5 points)
```{r}
# INSERT YOUR CODE HERE AND EXPLAIN YOUR ANSWER AS A COMMENT.

cor.test(USDAclean$Calories, USDAclean$TotalFat)

# The correlation between Calories and Total Fat is statistically significant
#as the p-value, 2.2e-16, is less than the confidence level of 0.05.

```

#### 18. Create a Linear Regression Model, using Calories as the dependent variable Protein, Total Fat, Carbohydrate, Sodium and Cholesterol as the independent variables. (5 points)
```{r}
# WRITE YOUR ANSWER AS A COMMENT.

lm_USD <- lm(USDAclean$Calories ~ USDAclean$Protein + USDAclean$TotalFat+
               USDAclean$Carbohydrate + USDAclean$Sodium + USDAclean$Cholesterol)
summary(lm_USD)

# y = 3.9882753 + 3.9891994(Protein) + 8.7716980(TotalFat) + 3.7432001(Carbohydrate) + 0.0003383(Sodium) + 0.0110138(Cholesterol)

```

#### 19. Which independent variable is the least significant? Why? (Hint: Use ANOVA) (5 points)
```{r}
# INSERT YOUR CODE HERE AND EXPLAIN YOUR ANSWER AS A COMMENT.

lmAnova_USD <- anova(lm_USD)

lmAnova_USD

# Sodium is the least significant variable. We can interpret this from the
#Analysis of Variance Table. The p-value for sodium is 0.1378, which is not
#very significant, especially compared to the p-values of the other variables,
#which are all much smaller, less than 2e-16.

```

#### 20. A new product is just produced with the following data: Protein=0.1, TotalFat=37, Carbohydrate=400, Sodium=440, Cholesterol=75, Sugar=NA, Calcium=35, Iron=NA, Potassium=35, VitaminC=10,	VitaminE=NA, VitaminD=NA. Based on the model you created, what is the predicted value for Calories? (5 points)
```{r}
# INSERT YOUR CODE HERE.



New_Product <- data.frame(Protein=0.1, TotalFat=37, Carbohydrate=400, Sodium=440, Cholesterol=75, Sugar=NA, Calcium=35, Iron=NA, Potassium=35, VitaminC=10,	VitaminE=NA, VitaminD=NA)


Predicted_Calories <- 3.9882753 + 3.9891994*New_Product$Protein +
8.7716980*New_Product$TotalFat + 3.7432001*New_Product$Carbohydrate +
0.0003383*New_Product$Sodium + 0.0110138*New_Product$Cholesterol

Predicted_Calories

```

#### 21. If the Sodium amount increases from 440 to 44440 (10000% increase), how much change will occur on Calories in percent? Explain why? (5 points)
```{r}
# INSERT YOUR CODE HERE AND EXPLAIN YOUR ANSWER AS A COMMENT.

Predicted_Calories_Increased <- 3.9882753 + 3.9891994*New_Product$Protein +
8.7716980*New_Product$TotalFat + 3.7432001*New_Product$Carbohydrate +
0.0003383*44440 + 0.0110138*New_Product$Cholesterol

Predicted_Calories_Increased

Change_in_Calories <- (44440-440)*0.0003383
Percentage_of_Change <- (Change_in_Calories/Predicted_Calories)* 100
Percentage_of_Change

# If the value of Sodium increased from 440 to 44440, the value of Calories
#would change by 14.8852. This represents a 0.81% change in the value of
#Calories from when Sodium was equal to 440. To get this result we multiply
#the difference in the Sodium value from before to after (44440-440 = 44000)
#by the coefficient for Sodium from the model, which is 0.0003383. We use this
#value in our regression calculation. The coefficient describes the change in
#the dependent variable for each unit of change in the Sodium variable.

```


This is the end of Assignment 1

Ceni Babaoglu, PhD
