---
title: "Multiple Regression"
output: html_document
---
we will have some examples on variable selection for multiple linear regression.
Preparton:
```{r}
install.packages("RCurl")
install.packages("MASS")
install.packages("leaps")

library(RCurl)
library(MASS)
library(leaps)

```


1.Create a multiple linear regression model to predict price.

In our examples, we will use the following dataset:
```{r}
c_price <- read.csv(file="http://vincentarelbundock.github.io/Rdatasets/csv/Ecdat/Computers.csv", sep = ",")
df <- data.frame(c_price)
str(df)
```

We will split the dataset to 70% of training and 30% of test sets. We want to make sure that the training set and the test set do not have any common data points.

```{r}
rn_index <- sample(1:nrow(c_price), floor(0.7 *nrow(c_price)))
rn_trian <- c_price[rn_index, ]
rn_test <- c_price[-rn_index, ]

```

Let's build a multiple linear regression model to predict the 'price' variable. Let's use “ram”, “screen”, “speed”, “hd” and “ads” as independent variables. We will train our model on the training set and do the prediction on the test set.

```{r}
rn_model <- lm(price~ram+speed+screen+hd+ads , data=rn_trian)
rn_prediction <- predict(rn_model , interval = "prediction", newdata = rn_test)

```

```{r}
rn_prediction
```

Let's see the errors and plot them on a histogram.

```{r}
errors <- rn_prediction[,"fit"]- rn_test$price
hist(errors)

```

Let's compute the root mean square error 

```{r}
rn_RMSE <-sqrt(mean(errors^2))
rn_RMSE

```

2.Apply forward selection and backward elimination algorithms to the dataset.

```{r}
Full_Model <- lm(price~ram+speed+screen+hd+ads+trend , df)
Null_Model <- lm(price~1 , df)

library(MASS)

sepF <- stepAIC(Null_Model, scope=list(lower=Null_Model, upper=Full_Model), direction="forward", trace=TRUE)
summary(sepF)
sepB <- stepAIC(Full_Model,  direction="backward", trace=TRUE)
summary(sepB)
setwi <-  stepAIC(Full_Model,  direction="both", trace=TRUE)

```

3.Apply regsubsets() to see the best combination of the attributes.

```{r}
library(leaps)

regsubset <- summary(regsubsets(price~ram+speed+screen+hd+ads+trend , df , nbest=1)) 

regsubset
regsubset$outmat
```


