---
title: "Lab8"
output: html_document
---
1. Apply PCA on the independent variables (first four columns) of the iris dataset by using
prcomp function from stats library

```{r}
data("iris")
str(iris)

iris_ind <- iris[,1:4]
library(stats)

iris_pca_rawdata <- prcomp(iris_ind, scale= FALSE , center = FALSE)
iris_pca_rawdata

iris_pca_rawdata$rotation




```
2. Get the transformed raw data.

```{r}
head(iris_pca_rawdata$x)


```

3. Plot the proportion of variance explained by each component. How many components will you choose to capture maximum variability in the dataset?


```{r}
plot(iris_pca_rawdata , type="l", main="without data normalization")


```
The heuristic rule is to choose components that form the elbow of the plot and drop the components after the elbow. So, we can drop components 3 and 4 without losing significant information.

4. Apply normalization (scaling and centering) to your data. Recalculate PCA with normalized data. What is the effect of normalizing on the PCA results?


```{r}

iris_pca_normdata <- prcomp(iris_ind , scale= TRUE , center = TRUE)
iris_pca_normdata
iris_pca_normdata$rotation
head(iris_pca_normdata$x)

plot(iris_pca_normdata , type="l", main="with data normalization")

```

5. Check the correlations of the original, transformed and normalized transformed data.


```{r}
# Correlations of the original data:
cor(iris_ind)

# Correlations of the transformed data
cor(iris_pca_rawdata$x)

# Correlation of the normalized data
cor(iris_pca_normdata$x)

```


