---
title: "Lab 9"
output: html_document
---
Preparation
```{r}
install.packages("class")
install.packages("gmodels")
library(class)
library(gmodels)
```



1. Download Prostate Cancer.csv from the course shell. The dataset contains patients
who have been diagnosed with either Malignant (M) or Benign (B) cancer

```{r}
Cancer_Dataset <- read.csv("C:/Users/Asma/Documents/prostate_cancer_dataset.csv", sep = "," , stringsAsFactors = FALSE)

head(Cancer_Dataset)

```
2. Check the data characteristics. Remove the first variable (id) from the dataset. Get the
numbers of patients in each class.

```{r}
str(Cancer_Dataset)
Cancer_Dataset <- Cancer_Dataset[-1]
table(Cancer_Dataset$diagnosis_result)
```

The variable diagnosis result is the target variable. This variable will determine the
results of the diagnosis based on the 8 numeric variables. In case you wish to rename B
as Benign and M as Malignant and see the results in the percentage form, you may
write:

```{r}
Cancer_Dataset$diagnosis_result<- factor(Cancer_Dataset$diagnosis_result,
levels = c("B", "M"),
labels = c("Benign", "Malignant"))

round(prop.table(table(Cancer_Dataset$diagnosis_result))*100, digits = 1)

```

3. Create a function to normalize a variable. Normalize the numeric variables in the dataset
to transfer their values to a common scale.

```{r}

Normalize <- function(x){
  return((x-min(x))/(max(x)-min(x)))
}

```

```{r}
Cancer_Dataset_N <- as.data.frame(lapply(Cancer_Dataset[2:9], Normalize))
Cancer_Dataset_N <-cbind(Cancer_Dataset$diagnosis_result, Cancer_Dataset_N)

```
The first variable in the dataset (after removal of id) is diagnosis result which is not
numeric in nature. So, you start from the second variable. The function lapply()
applies normalize() to each feature in the data frame.

Using the variable radius, you can check whether the data has been normalized

```{r}
summary(Cancer_Dataset_N$radius)
```
4. Split the dataset into training and test sets.

```{r}
set.seed(123)

Cancer_Dataset_Index <- sample(1:nrow(Cancer_Dataset_N), 0.65* nrow(Cancer_Dataset_N))

Cancer_Dataset_train <- Cancer_Dataset_N[Cancer_Dataset_Index, ]
Cancer_Dataset_test <- Cancer_Dataset_N[-Cancer_Dataset_Index,]

```

5. Use the KNN algorithm to predict the cancer

```{r}
Cancer_train_labels <- Cancer_Dataset_train[,1]
Cancer_test_labels <- Cancer_Dataset_test[,1]


```

```{r}

Cancer_test_Pred <- knn(train=Cancer_Dataset_train[,2:9], test=Cancer_Dataset_test[,2:9], cl=Cancer_Dataset_train[,1], k=10)

```
6. Evaluate the model performance

```{r}
table(Actual=Cancer_test_labels , Predicted = Cancer_test_Pred)

```

